Quality of Service with Software Defined Network.

### 2017-02-16 report/paper.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \textit{\textbf{SDN}, Software Defined Network} is a new approach to create a network of computer and networks, which is based on the
most famous OpenFlow protocol. The key concepts of this type of networks are the hierarchy, the flow control and the statistics.
The hierarchy allows a network administrator to set up a controller which will choose how to make the network evolve setting up
how the routing will switch the traffic and doing so, how the single switch will behave.
The flow control is done by the controller that, running the algorithm the network administrator created, choose the path for a
type of traffic that occurs between a source and a destinaion host.
The most important and interesting thing about those type of networks is the statistics: every switching unit collects data from its
interfaces (e.g. the number of packets going in or out through a specific port or interface). This allow the controller to instantly
how the network is evolving and, if necessary, switch a traffic from congestioned links to other that, in that moment, are almost fully loaded.
Exploiting the statistic capabilities is the key to manage wisely a network and, since there aren't many papers about these potentialities,
we decided to focus on that. In this paper we will show how we handled the contemporary traffic between multiple host, both inside SDN network
nd between the SDN one and the external internet. Starting from traffic analysis and going through the SDN state of the art, we implemented some function
to exploit better the SDN POX controller and we reach the end of this paper showing how we found three different types of traffic and
for each one, we provided different metrics and different path to optimize the load. We realized the controller using POX, a python-based controller
and we tested it with the switches that were provided us in the Communication Networks Lab at DEI.
\subsection{Known Limitations}
SDN networks are known to be underperforming with low traffic, which means that if nodes tries to connect with each other
to exchange little traffic, the network will have a higher delay and will perform worst than a
traditional network. We tried to reduce this underperformance with the usage of the wildcards:
Every rule needs at least the source ip or the destination ip. If one of them is in a network
(e.g. the internet traffic is in the network 0.0.0.0/0 excluded 192.168.10.0/24 ),
we can redirect all the traffic from an internal host to the internet with a predefined
path. This means that only one PacketIn will be sent to the controller if a host is surfing the network.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec:sota}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A good introduction about the SDN can be the  "Simulation in an SDN network scenario using the
POX Controller" where has been presented the SDN environment and the mininet simulation
tool that creates a topology where to simulate the switches, the traffic between hosts and
allow the controller to communicate with the switches. The work presented in
"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad [put cit]
has been implemented to improve the controller speed and efficiency as it make it more reactive.
The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found
in the paper called "Getting traffic statistics from network devices in an
SDN environment using OpenFlow" by D.J. Hamad & others. [ put note here]
For traffic analysis we haven't found any paper describing the characteristics of
some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section \ref{sec:symo}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Model}\label{sec:symo}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The System model is a description of your operating assumptions with related motivation and justification

@@

\subsection{Traffic analysis}\label{sec:traf}
We used laptop connected via
GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
No other traffic sources were active during the testing, so the router was handling only
the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
#Mb for the gaming traffic, #Mb for the VoIP traffic and 3GB for the torrent traffic.
The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)

The traffic analysis of the torrent protocol shows that the average packet size is the MTU
(i.e. 1500bytes) and the protocol tryies to exploit all the channel capabilities. The traffic
is made via UDP packets with the port range in the ephimeral ports (2000_- 65535).
Moreover, the node is linked to a lot of IPs with multiple connection: this means that
if someone wants to find a p2p traffic, it can look to the statistics and see if the traffic is loading all the channel,
to how many IPs, an host is connected to and if it uses UDP.
For the VoIP we saw that the protocol uses UDP and the traffic, after a transition time, goes only between two hosts.
The average packet size is the MTU and the protocol tryies to exploit all the channel for a better QoE.
On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
decreased to allow communication. The traffic is between the UDP port 80 ##TODO## TOCHECK ##

The last traffic analysis we made was with a DOTA[explain what it is] web match.
We saw that for that type of gaming session, the packet size is about # bytes and involves both TCP and UDP ports.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \textit{\textbf{SDN}, Software Defined Network} is a new approach to create a network of computer and networks, which is based on the
most famous OpenFlow protocol. The key concepts of this type of networks are the hierarchy, the flow control and the statistics.
The hierarchy allows a network administrator to set up a controller which will choose how to make the network evolve setting up
how the routing will switch the traffic and doing so, how the single switch will behave.
The flow control is done by the controller that, running the algorithm the network administrator created, choose the path for a
type of traffic that occurs between a source and a destinaion host.
The most important and interesting thing about those type of networks is the statistics: every switching unit collects data from its
interfaces (e.g. the number of packets going in or out through a specific port or interface). This allow the controller to instantly
how the network is evolving and, if necessary, switch a traffic from congestioned links to other that, in that moment, are almost fully loaded.
Exploiting the statistic capabilities is the key to manage wisely a network and, since there aren't many papers about these potentialities,
we decided to focus on that. In this paper we will show how we handled the contemporary traffic between multiple host, both inside SDN network
and between the SDN one and the external internet. Starting from traffic analysis and going through the SDN state of the art, we implemented some function
to exploit better the SDN POX controller and we reach the end of this paper showing how we found three different types of traffic and
for each one, we provided different metrics and different path to optimize the load. We realized the controller using POX, a python-based controller
and we tested it with the switches that were provided us in the Communication Networks Lab at DEI.
\subsection{Known Limitations}
SDN networks are known to be underperforming with low traffic, which means that if nodes tries to connect with each other
to exchange little traffic, the network will have a higher delay and will perform worst than a
traditional network. We tried to reduce this underperformance with the usage of the wildcards:
Every rule needs at least the source ip or the destination ip. If one of them is in a network
(e.g. the internet traffic is in the network 0.0.0.0/0 excluded 192.168.10.0/24 ),
we can redirect all the traffic from an internal host to the internet with a predefined
path. This means that only one PacketIn will be sent to the controller if a host is surfing the network.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec:sota}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A good introduction about the SDN can be the  "Simulation in an SDN network scenario using the
POX Controller" where has been presented the SDN environment and the mininet simulation
tool that creates a topology where to simulate the switches, the traffic between hosts and
allow the controller to communicate with the switches. The work presented in
"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad [put cit]
has been implemented to improve the controller speed and efficiency as it make it more reactive.
The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found
in the paper called "Getting traffic statistics from network devices in an
SDN environment using OpenFlow" by D.J. Hamad \& others. [ put note here]
For traffic analysis we haven't found any paper describing the characteristics of
some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section \ref{sec:symo}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%parte scritta da Matteo da tradure e forse da spostare
%\section{Controller Implementation}\label{sec:Implementation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%We developed a Pox based controller using Python2.
%The main function that we need and that we had developed are:
%\begin{itemize}
% \item Gloal knowledg of our network topology with link performance metrics
% \item Ensure host to host connectivity and external networ reachability using different performance path
% \item Proces packetIn request
%\end{itemize}

%\subsection{Network topology}
%To storage our network topology we use a specific python library called \emph{NetworkX} this library 
%\url{https://networkx.github.io/}
%give us the possibility
%to create a graph. We have create this graph with two types of nodes: Switch and Hosts, all of them connected with link.
%For now we use a single-edge graph but a real application have to use double-edge graph in orther to measure both link's performances.
%\newline There is a principal function implemented by Pox called dyscovery.py that give us the event regarding ( check ) host, switch
%and link connection/disconnection so handling this event we can build our graph step-by-step.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Model}\label{sec:symo}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
(The System model is a description of your operating assumptions with related motivation and justification)

@@

\subsection{Traffic analysis}\label{sec:traf}
We used laptop connected via
GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
No other traffic sources were active during the testing, so the router was handling only
the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
\#Mb for the gaming traffic, \#Mb for the VoIP traffic and 3GB for the torrent traffic.
The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)

The traffic analysis of the torrent protocol shows that the average packet size is the MTU
(i.e. 1500bytes) and the protocol tryies to exploit all the channel capabilities. The traffic
is made via UDP packets with the port range in the ephimeral ports (2000_- 65535).
Moreover, the node is linked to a lot of IPs with multiple connection: this means that
if someone wants to find a p2p traffic, it can look to the statistics and see if the traffic is loading all the channel,
to how many IPs, an host is connected to and if it uses UDP.
For the VoIP we saw that the protocol uses UDP and the traffic, after a transition time, goes only between two hosts.
The average packet size is the MTU and the protocol tryies to exploit all the channel for a better QoE.
On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
decreased to allow communication. The traffic is between the UDP port 80 \#\#TODO\#\# TOCHECK \#\#

The last traffic analysis we made was with a DOTA[explain what it is] web match.
We saw that for that type of gaming session, the packet size is about \# bytes and involves both TCP and UDP ports.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%da vedere come distribuire
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{matteo}

\section{Program structure}

On our controller in orther two offer principals function we have two storage two principal things: the logical topology of the SDN network and the 
proprirty of every nodes.

\subsection{Topology}
E' necessario memorizzare la topologia della rete in una struttura dati che permetta la facile ricerca di cammini minimi tra due host. 
Abbiamo utilizzato quindi una struttura a grafo semplice in cui i nodi interni sono gli switch e le foglie sono gli host della nostra rete.
Nella realtà le connessioni logiche tra due switch possono prevedere più di un link fisico che può essere mono o bdirezionale.
Nel nostro sistema abbiamo preferito rappresentare la connessione tra due switch con un unico edge bidirezionale, da cui poi è facilmente estendibile
in vista di un' applicazione pratica.

Per gestire il grafo viene utilizzata la libreria scritta in python NeworkX

The scope of our structure is storage the logically connection of network's nodes in orther to quikly discovery the minimum path

\subsection{Network knowledge}\label{sec:topo}
Our controller program need to know SDN all network topology in orther to guided ( instradare ) the packet along a better path or along a path to ensure 
a better ( funzionamento ) on SDN network.
We use two type of date structur:
\begin{itemize}
 \item \emph{Switch and Node}, we create ad-hoc object to save the main information of every single Node and Switch. Inside switch opject there are many 
 dictionary that ( permettono ) to find quikly port to switch association, ip to port association and other. [da sistemare con piu' precisione]
 \item \emph{Logically connection} to have a complete knowledg of the topology we use graph structured, despitly ( diversamente ) reality in
 whitch  double-edge used we use single-edge graph to avoid evil implementation difficult that esule our aim. The NetworkX python library give as method
 to create a graph in witch every node is a pointer to real Host or Switch object and the link conceptually have various metrics attribute.
 ``Conceptually'' because NetworkX library don't give us the possibility to save more than one single weight attribute for each link, so 
 we need to mantain multy graph, each of every metrics used with the same topology atribute but different weigth on theyr links.
\end{itemize}
